\chapter{Lecture: Binary Sort, Insert Sort, Merge Sort}
\href{https://ocw.mit.edu/courses/electrical-engineering-and-computer-science/6-006-introduction-to-algorithms-fall-2011/lecture-videos/lecture-3-insertion-sort-merge-sort/}{Lecture 3} from the MIT's 
Intro to Algorithms.

Why sorting? Easy to search like phone books, mp4 organizations, etc. Problems can become easy
once items are sorted, example, finding a median.
\begin{python}
def binary_search_recursive(array, element, start, end):
    if start > end:
        return -1

    mid = (start + end) // 2
    if element == array[mid]:
        return mid

    if element < array[mid]:
        return binary_search_recursive(array, element, start, mid - 1)
    else:
        return binary_search_recursive(array, element, mid + 1, end)
        
def binary_search_iterative(array, element):
    mid = 0
    start = 0
    end = len(array)

    while (start <= end):
        mid = (start + end) // 2

        if element == array[mid]:
            return mid

        if element < array[mid]:
            end = mid - 1
        else:
            start = mid + 1
    return -1
    
def insertion_sort(array):
    # We start from 1 since the first element is trivially sorted
    for index in range(1, len(array)):
        currentValue = array[index]
        currentPosition = index

        while (currentPosition > 0 and 
            array[currentPosition - 1] > currentValue):
            array[currentPosition] = array[currentPosition -1]
            currentPosition = currentPosition - 1

        array[currentPosition] = currentValue
\end{python}
Data compression benefits from sorting. Insertion sort, works by doing pair-wise swaps. What is the time
complexity? We have \(\theta(n)\) steps but each step could do \(\theta(n)\) work (swaps and compares)
 so \(\theta(n^2)\). 
 
 Let's consider Merge Sort, another divide and conquer algorithm. Split array into left and right halves.
 Then we merge and sort array. However, with merge sort, we need \(\theta(n)\) auxiliary space whereas
 with insertion sort, we need \(\theta(1)\) auxiliary space.
 \begin{python}
 def merge_sort(list):
    list_length = len(list)

    if list_length == 1:
        return list

    mid_point = list_length // 2

    left_partition = merge_sort(list[:mid_point])
    right_partition = merge_sort(list[mid_point:])

    return merge(left_partition, right_partition)

def merge(left, right):
    output = []
    i = j = 0

    while i < len(left) and j < len(right):
        if left[i] < right[j]:
            output.append(left[i])
            i += 1
        else:
            output.append(right[j])
            j += 1

    output.extend(left[i:])
    output.extend(right[j:])

    return output
 \end{python}







